File-based localization of numerical perturbations in data analysis pipelines

### 2020-04-06 paper.tex
Update Fig4

We processed the 20 subjects with PreFreeSurfer and FreeSurfer, using the 2
CentOS versions. Each subject was processed twice on the same operating
system to detect within-OS variability coming from pseudo-random
operations. We ignored execution-specific information such as file path or
timestamps by using FreeSurfer tools \texttt{mri\_diff},
\texttt{mris\_diff}, and \texttt{lta\_diff} that determine if MRI volumes
or transformations differ. To compare segmentations, we used the Dice coefficient defined as follows:
\[DICE=\frac{2|X \cap Y|}{|X| + |Y|}\]

@@

Figure~\ref{fig:pfs_freq} shows the frequency of non-reproducible pipeline processes
in PreFreeSurfer. 
Differences were observed in linear registration 
with FSL FLIRT (in ACPC-Alignment, Brain Extraction, Distortion Correction, and
Atlas Registration), in non-linear registration with FSL FNIRT (in Brain Extraction 
and Atlas Registration), and in image warping with FSL \texttt{new\_invwarp} (in Brain Extraction 
and Atlas Registration). Differences were also observed in image mean 
computations with FSL maths  (in Anatomical Average). 
A complete provenance graph of the PreFreeSurfer is illustrated in Figure~\ref{fig:complete_pfs}
\begin{figure}
  \centering
    \includegraphics[width=\columnwidth]{images/pfs-intransient} 
    \caption{A complete provenance graph from the PreFreesurfer pipeline.
     Full-resolution image is available 
     \href{https://drive.google.com/file/d/1xonnMq4hLEe6GwMp4P7umikvrIAeflDe/view?usp=sharing}{here}.}
    \label{fig:complete_pfs}
\end{figure}
% Between subject differences can be a result of different subject types. 
% For instance, the subjects contain two T1 images need more calculations 
% to produce an average image between them.

@@

While by design \toolname only identifies processes that create differences
between conditions, pipeline results may still differ due to the
propagation and amplification of differences created at previous processing
steps, PreFreeSurfer in our case. We observed the effect of this
propagation in FreeSurfer results, as shown in
Figure~\ref{fig:tissue_class} for whole-brain segmentations. Furthermore,
the Dice coefficients associated with the 44 regions segmented by
FreeSurfer are shown in Figure~\ref{fig:scatter_plot}, showing, for some subjects, Dice values
below 0.9 in structures such as the corpus callosum, pallidum, amygdala,
accumbens area, and ventricles.

@@

\begin{figure}
  \hspace*{-1cm}
  \includegraphics[width=1.2\linewidth]{images/dice_regions.pdf}
    \caption{Dice coefficients between regions segmented by FreeSurfer in CentOS6 vs CentOS7 (N=20).}
    \label{fig:scatter_plot}
  \end{figure}

>>>>

We processed the 20 subjects with PreFreeSurfer and FreeSurfer, using the 2
CentOS versions. Each subject was processed twice on the same operating
system to detect within-OS variability coming from pseudo-random
operations. We ignored execution-specific information such as file path or
timestamps by using FreeSurfer tools \texttt{mri\_diff},
\texttt{mris\_diff}, and \texttt{lta\_diff} that determine if MRI volumes
or transformations differ. To compare segmentations, we used the Dice coefficient defined as follows:
\[DICE=\frac{2|X \cap Y|}{|X| + |Y|}\].

@@

Figure~\ref{fig:pfs_freq} shows the frequency of non-reproducible pipeline processes
in PreFreeSurfer. 
Differences were observed in linear registration 
with FSL FLIRT (in ACPC-Alignment, Brain Extraction, Distortion Correction, and
Atlas Registration), in non-linear registration with FSL FNIRT (in Brain Extraction 
and Atlas Registration), and in image warping with FSL \texttt{new\_invwarp} (in Brain Extraction 
and Atlas Registration). Differences were also observed in image mean 
computations with FSL maths  (in Anatomical Average). 
A complete provenance graph of the PreFreeSurfer is illustrated in Figure~\ref{fig:complete_pfs}
\begin{figure}
  \centering
    \includegraphics[width=\columnwidth]{images/pfs-labeled.png} 
    \caption{A complete provenance graph from the PreFreesurfer pipeline.
     Full-resolution image is available 
     \href{https://drive.google.com/file/d/1BKq5ZXIVr3bkij88SmeW7ulWOExSCm7n/view?usp=sharing}{here}.}
} 
    \label{fig:complete_pfs}
\end{figure}
% Between subject differences can be a result of different subject types. 
% For instance, the subjects contain two T1 images need more calculations 
% to produce an average image between them.

@@

While by design \toolname only identifies processes that create
differences between conditions, pipeline results may still differ due to
the propagation and amplification of differences created at previous
processing steps, PreFreeSurfer in our case. We observed the effect of this
propagation in FreeSurfer results, as shown in
Figure~\ref{fig:tissue_class} for whole-brain segmentations. Furthermore,
the Dice coefficients associated with the 44 regions segmented by
FreeSurfer are shown in Figure~\ref{fig:scatter_plot}, showing low Dice
values (below 0.5) in the smallest structures.

@@

\begin{figure*}
  \centering
  \begin{subfigure}{0.45\linewidth}
    \includegraphics[width=\linewidth]{images/scatter_plot_1.png}
    \caption{Regions with \emph{less} than 10,000 voxels.}
  \end{subfigure}
  \hfill
  \begin{subfigure}{0.45\linewidth}
    \includegraphics[width=\linewidth]{images/scatter_plot_2.png}
    \caption{Regions with \emph{more} than 10,000 voxels.}
  \end{subfigure}
    % \includegraphics[width=\textwidth]{images/scatter_plot.png} 
    \caption{Dice coefficients between regions segmented by FreeSurfer in CentOS6 vs CentOS7 (N=20).
        To improve readability, the two graphs are on different y scales.}
    \label{fig:scatter_plot}
  \end{figure*}
